\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-05-20 manuscript.tex
Formatting

\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}

@@

@@

The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte. Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in domesticated maize. Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication, led us to investigate its prevalence and potential role in teosinte. Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp. \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mexicana}. Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte.  Population genetic signatures, however, are consistent with selection on this locus and suggest the \emph{Hopscotch} insertion at \emph{tb1} may play an ecological role in teosinte. Finally, two greenhouse experiments with \emph{Zea mays} ssp. \emph{parviglumis} do not suggest \emph{tb1} controls tillering in natural populations of this subspecies. Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte differs from domesticated maize, and that other loci may play a role in observed variation in this trait.

>>>>

\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}

@@

\emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte. Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in domesticated maize. Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication, led us to investigate its prevalence and potential role in teosinte.

\vspace*{0.2cm}

\emph{Methods:}

\vspace*{0.2cm}

\emph{Key results:} Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp. \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mexicana}. Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte.  Population genetic signatures, however, are consistent with selection on this locus and suggest the \emph{Hopscotch} insertion at \emph{tb1} may play an ecological role in teosinte. Finally, two greenhouse experiments with \emph{Zea mays} ssp. \emph{parviglumis} do not suggest \emph{tb1} controls tillering in natural populations of this subspecies.

\vspace*{0.2cm}

\emph{Conclusions:} Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte differs from domesticated maize, and that other loci may play a role in observed variation in this trait.

\vspace*{0.2cm}

@@
